Sequential algorithms to split and merge ultra-high resolution 3D images

### 2017-07-24 paper.tex
Updated figure 2 to reflect the new cases
Fixed figure 2 (again)

The buffer is a
slightly complex data structure, for instance an associative array or
a Python dictionary, capable of storing multiple disjoint sequences of
contiguous bytes without having to allocate memory for the bytes
between such sequences. Figure~\ref{fig:cluster-reads-buffer}
illustrates how the buffer would fill up when two blocks of one slice
and three rows per slice are loaded. Implementation details about the
buffer are given later.
\begin{figure}
\centering
\def\svgwidth{\columnwidth}
\input{figures/buffer-cluster-reads.pdf_tex}
\caption{Illustration of the buffer used in cluster reads. White
  portions are not allocated. [a,b] refers to block a, row b.}
\label{fig:cluster-reads-buffer}
\end{figure}

The number of seeks performed by cluster reads depends on how blocks
loaded in memory arrange in the reconstructed image. In the best case,
complete contiguous slices of the reconstructed image can be assembled
in memory and written without any seeking. In the worst case, some
rows in the reconstructed image are not covered by blocks loaded in
memory: O($d^2$) seeks are then required to write the memory load, one
for every partial row in every partial slice. In the intermediary
case, rows are complete but some slices can only be partially
reconstructed: O($d$) seeks are then
required. In the general case, we have:
\begin{equation}
  N_\mathrm{cluster\_reads}= n + \sum_{i=1}^{\ceil*{\frac{bR}{m}}}f_i
\end{equation}
where $f_i$ is the number of seeks done in the $i^{th}$ memory
load. Figure~\ref{fig:cluster-reads} lists the possible block
arrangements and the corresponding values of $f_i$. \todo{Val\'erie:
  Check that ordering in the Figure is consistent with Nifti's.}  From
this first analysis, we conjecture that incomplete block rows and
incomplete block slices should be avoided whenever possible. That is,
our implementation of cluster reads will consider only three cases:
\begin{itemize}
  \item Case 1: $m<d^3\sqrt[3]{n}b=\frac{D^3}{\sqrt[3]{n}^2b}$, that is, a
    row of blocks does not fit in memory. Incomplete rows of
    blocks will be merged, with complexity O($d^2$).
  \item Case 2: $\frac{D^3}{\sqrt[3]{n}^2}b\leq m<\frac{D^3}{\sqrt[3]{n}}b$,
    that is, a row of blocks fits in memory but a slice of blocks does
    not. Only complete rows of blocks will be merged, with
    complexity O(d). We have $m'=\floor*{\frac{m}{d^3\sqrt[3]{n}b}}$.
    \item Case 3: $\frac{D^3}{\sqrt[3]{n}}b \leq m < D^3b$, that is, a slice
      of blocks fits in memory. Only complete slices of blocks
      will be merged, with complexity O(1). We have
      $m'=\floor*{\frac{m}{d^3\sqrt[3]{n}^2b}}$.
\end{itemize}

@@

\begin{figure}
\centering
\def\svgwidth{0.3\columnwidth}
\input{figures/case1.pdf_tex}
\def\svgwidth{0.3\columnwidth}
\input{figures/case2.pdf_tex}
\def\svgwidth{0.3\columnwidth}
\input{figures/case3-a.pdf_tex}
\def\svgwidth{0.3\columnwidth}
\input{figures/case3-b.pdf_tex}
\def\svgwidth{0.3\columnwidth}
\input{figures/case4.pdf_tex}
\def\svgwidth{0.3\columnwidth}
\input{figures/case5.pdf_tex}
\def\svgwidth{0.3\columnwidth}
\input{figures/case6-a.pdf_tex}
\def\svgwidth{0.3\columnwidth}
\input{figures/case6-b.pdf_tex}
\def\svgwidth{0.3\columnwidth}
\input{figures/case7.pdf_tex}

\caption{In-memory block arrangements leading to different number of
  seeks in cluster reads. Blue blocks need seeking before each of
  their slices ($d$ seeks in total). Red blocks need seeking before
  each of their rows ($d^2$ seeks). Purple blocks require seeking
  before each row and each slice ($d^2$ seeks). Green blocks are
  written with no seeking.}
\label{fig:cluster-reads}
% See source files in
% https://drive.google.com/drive/folders/0BzT7pbvPDUdQRXppVlhfdS0ycGc?usp=sharing
\end{figure}

The number of seeks performed by cluster reads depends on the three
cases listed above. In case 1, $f_i$ might be (a) $d^2$
 or (b) $2d^2$
depending whether the memory load
overlaps on multiple block slices, which will happen
$\alpha$
times where:
\begin{equation*}
  \alpha=\sqrt[3]{n}-1-\floor*{\frac{n-\sqrt[3]{n}^2}{LCM\left(\frac{m'}{d^3},\sqrt[3]{n}^2\right)}}
\end{equation*}
 We have:
\begin{equation}
  N_\mathrm{cluster\_reads\_v1}= n + \frac{bR}{m'}d^2+d^2\alpha
\end{equation}
Similarly, in Case 2 we have:
\begin{equation}
  N_\mathrm{cluster\_reads\_v2}= n + \frac{bR}{m'}d+d\alpha
\end{equation}
And in Case 3:
\begin{equation}
  N_\mathrm{cluster\_reads\_v3}= n + \frac{bR}{m'}
\end{equation}

>>>>

The buffer might be a
slightly complex data structure, for instance an associative array or
a Python dictionary, capable of storing multiple disjoint sequences of
contiguous bytes without having to allocate memory for the bytes
between such sequences. Figure~\ref{fig:cluster-reads-buffer}
illustrates how the buffer would fill up when two blocks of one slice
and three rows per slice are loaded. Implementation details about the
buffer are given later.
\begin{figure}
\centering
\def\svgwidth{\columnwidth}
\input{figures/buffer-cluster-reads.pdf_tex}
\caption{Illustration of the buffer used in cluster reads. White
  portions are not allocated. [a,b] refers to block a, row b.}
\label{fig:cluster-reads-buffer}
\end{figure}

The number of seeks performed by cluster reads depends on how blocks
loaded in memory arrange in the reconstructed image. In the best case,
complete contiguous slices of the reconstructed image can be assembled
in memory and written without any seeking. In the worst case, the
memory load only partially covers rows in the reconstructed image:
O($d^2$) seeks are then required during writing, one for every partial
row in every partial slice. In the intermediary case, rows are
complete but some slices can only be partially reconstructed: O($d$)
seeks are then required.

\todo{Check that ordering in the Figure is consistent with Nifti's.}

Incomplete block rows and incomplete block slices lead to extra
seeking and should be avoided as much as possible.  Thus, cluster
reads consider only three cases, represented in
Figure~\ref{fig:cluster-reads}:
\begin{itemize}
  \item Case 1: $m<d^3\sqrt[3]{n}b=\frac{D^3}{\sqrt[3]{n}^2b}$, that is, a
    row of blocks does not fit in memory. Incomplete rows of
    blocks will be merged, with $d^2$ seeks per memory load.
  \item Case 2: $\frac{D^3}{\sqrt[3]{n}^2}b\leq m<\frac{D^3}{\sqrt[3]{n}}b$,
    that is, a row of blocks fits in memory but a slice of blocks does
    not. Only complete rows of blocks will be merged, with
    $d$ seeks per memory load. We have $m'=\floor*{\frac{m}{d^3\sqrt[3]{n}b}}$.
    \item Case 3: $\frac{D^3}{\sqrt[3]{n}}b \leq m < D^3b$, that is, a slice
      of blocks fits in memory. Only complete slices of blocks
      will be merged, a single seek per memory load. We have
      $m'=\floor*{\frac{m}{d^3\sqrt[3]{n}^2b}}$.
\end{itemize}

@@

\begin{figure}
\centering
\def\svgwidth{0.3\columnwidth}
\input{figures/case1-a.pdf_tex}
\def\svgwidth{0.3\columnwidth}
\input{figures/case2-a.pdf_tex}
\def\svgwidth{0.3\columnwidth}
\input{figures/case3.pdf_tex}

\caption{Memory-load configurations in cluster reads, leading to
  different number of seeks. Red blocks need seeking before each of
  their rows ($d^2$ seeks). Blue blocks need seeking before each of
  their slices ($d$ seeks in total). Green blocks need only a single
  seek. Grey, dashed, transparent blocks represent the contiguous
  memory loads and are added for the sake of visualization.}
\label{fig:cluster-reads}
% See source files in
% https://drive.google.com/drive/folders/0BzT7pbvPDUdQRXppVlhfdS0ycGc?usp=sharing
\end{figure}

The total number of seeks performed by cluster reads depends on the three
cases listed above and on the overall number of memory loads.
\begin{equation}
  N_\mathrm{cluster\_reads\_v1}= n + \frac{bR}{m'}d^2+d^2\alpha
\end{equation}
Similarly, in Case 2 we have:
\begin{equation}
  N_\mathrm{cluster\_reads\_v2}= n + \frac{bR}{m'}d+d\alpha
\end{equation}
And in Case 3:
\begin{equation}
  N_\mathrm{cluster\_reads\_v3}= n + \frac{bR}{m'}
\end{equation}
